---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DIMPLE: **D**istance based **I**nference for **M**ultip**L**ex imaging **E**xperiments


<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/DIMPLE)](https://CRAN.R-project.org/package=DIMPLE)
[![codecov](https://codecov.io/gh/nateosher/DIMPLE/branch/master/graph/badge.svg?token=87UMWJ42XU)](https://codecov.io/gh/nateosher/DIMPLE)
<!-- badges: end -->

The goal of DIMPLE is to provide tools and infrastructure to facilitate the manipulation, exploration, and modeling of multiplex imaging data.

## Installation

You can install the development version of DIMPLE from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("nateosher/DIMPLE")
```

To install with vignettes (which we recommend), use the command:
``` r
# install.packages("devtools")
devtools::install_github("nateosher/DIMPLE", vuild_vignettes = TRUE)
```

## Example

Suppose we're interested in exploring the lung cancer data from the [`VectraPolarisData`](https://bioconductor.org/packages/release/data/experiment/html/VectraPolarisData.html) package. Assuming you've already installed that package, we can build a `MltplxExperiment` object, the core of `DIMPLE`, from the data as follows:

```{r}
library(DIMPLE)
library(tidyverse)
lung_data = VectraPolarisData::HumanLungCancerV3()

# Get the x coordinates for the cells
cell_x_coords = SpatialExperiment::spatialCoords(lung_data)[,1] %>% as.numeric()
# Get the y coordinates for the cells
cell_y_coords = SpatialExperiment::spatialCoords(lung_data)[,2] %>% as.numeric()
# Get the patient ids for the slides
patient_ids = lung_data$slide_id
# Get the slide ids for the slides
slide_ids = lung_data$sample_id

# Map phenotypes to character vector - there are more efficient ways to do this,
# but this way is the most legible
test_n = length(lung_data$phenotype_cd14)
tictoc::tic()
cell_marks = tibble(
  cd_14 = lung_data$phenotype_cd14[1:test_n] == "CD14+",
  cd_19 = lung_data$phenotype_cd19[1:test_n] == "CD19+",
  cd_4 = lung_data$phenotype_cd4[1:test_n] == "CD4+",
  cd_8 = lung_data$phenotype_cd8[1:test_n] == "CD8+",
  ck = lung_data$phenotype_ck[1:test_n] == "CK+",
  other = lung_data$phenotype_other[1:test_n] == "Other+"
) %>%
  as.matrix() %>%
  (\(m){
    m %*% 1:6
  }) %>%
  map_chr(\(x) c(NA, "CD14", "CD19", "CD4", "CD8", "CK", "Other")[x + 1])

na_marks = which(is.na(cell_marks))

na_marks = which(is.na(cell_marks))
cell_x_coords = cell_x_coords[-na_marks]
cell_y_coords = cell_y_coords[-na_marks]
patient_ids = patient_ids[-na_marks]
slide_ids = slide_ids[-na_marks]
cell_marks = cell_marks[-na_marks]

lung_experiment = DIMPLE::new_MltplxExperiment(
  x = cell_x_coords,
  y = cell_y_coords,
  marks = cell_marks,
  slide_id = slide_ids
)

lung_experiment
```

This object can be indexed like a list- each entry is of class `MltplxObject`, and stores data about a particular biopsy:

```{r }
lung_experiment[[1]]
```

Using the implemented generic functions, this allows for quick inspection of a given biopsy:

```{r example_plot}
plot(lung_experiment[[1]])
```

To learn more about these data structures and how to utilize them for exploration, check out the `mltplx-experiment` vignette. `DIMPLE` also provides functions for simulating Multiplex Imaging data- an overview of this functionality can be found in the `simulations` vignette






















