---
title: "Example Analysis Workflow"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(DIMPLE)
library(tidyverse)
```

## Setting up the data

In this vignette, we will show an example of an analysis workflow that could be performed using the `DIMPLE` package.
We will be using a simulated point pattern dataset, along with simulated patient metadata that has a continuous `age` covariate and categorical `group` covariate. We will demonstrate how to analyze this data using the JSD distance metric. The below steps for seting up the `MltplxExperiment` object are similar to those performed in other vignettes, so we won't explain this process again here. 

```{r}
# Create an experiment with intensities, distance matrices and patient metadata
set.seed(1234)
N <- 30000
n_slides <- 30
cell_x_values = runif(N, 0, 600)
cell_y_values = runif(N, 0, 600)
cell_marks = sample(c("Tumor", "Immune", "Stroma","Other"), N, replace = TRUE)
slide_ids = rep(paste("Slide", 1:n_slides), each = 1000)

slides <- unique(slide_ids)
n_patients <- 10
n_groups <- 2
patient_ids <- rep_len(paste0("P",1:n_patients),n_slides)

metadata <- tibble::tibble(slide_id=slides,
                   patient_id=patient_ids)

groups <- sample(rep_len(sample(paste0("G",1:n_groups)),n_patients))

ages <- runif(n_patients,min=45,max=100)

gp_tb <- tibble::tibble(patient_id=unique(patient_ids),group=groups,age=ages)
metadata <- dplyr::left_join(metadata,gp_tb,by = "patient_id")
print(metadata)

mltplx_experiment = new_MltplxExperiment(x = cell_x_values,
                                  y = cell_y_values,
                                  marks = factor(cell_marks),
                                  slide_id = slide_ids,
                                  ps = 30, bw = 40,
                                  dist_metric = jsd,
                                  metadata = metadata)

```

Next we can add information on how the JSD between cell types changes in different areas of higher and lower tumor density, by using the `update_qdist` function. This function takes in the `MltplxExperiment` object, the distance metric of choice, a cell type by which the image is partitioned on (by the density of that cell type), as well as the quantiles that the mask density should be partitioned into.

```{r}
q_probs <- tibble(from=c(0,34,67),to=c(33,66,100))
mltplx_experiment <- update_qdist(mltplx_experiment,
                                  jsd,
                                  "Tumor",
                                  q_probs)

mltplx_experiment
```

## Analysis

Now we can start our analyses! First, we can simply plot some of the point pattern datasets, using `plot_ppp`.

```{r}
mltplx_experiment %>%
  plot_ppp(slide_ids = c("Slide 12","Slide 13"))
```

We can also plot the smoothed intensity surfaces, using `plot_intensity_surface`.

```{r}
mltplx_experiment %>%
  plot_intensity_surface(types = c("Tumor","Stroma"),
                   slide_ids = c("Slide 7","Slide 9"))

```

If the `MltplxExperiment` has distance matrices calculated, as well as multiple images per patient, we can look at how the distance between two cell types is distributed within each patient using `patient_boxplots`. This function also tests that the mean of the distribution for each patient is significantly different than a value `mu0` (default zero).

```{r}
mltplx_experiment %>%
  patient_boxplots(t1="Tumor",t2="Immune",grouping_var = "group",p_val_col = "p.adj")
```

We can also look at the distribution of the distance between two cell types, but now stratified into patient groups (rather than individual patients), by using the `plot_dist_boxplots` function. The image-level data is aggregated into patient-level data by a function `agg_fun`.
```{r}
mltplx_experiment %>%
  plot_dist_boxplots(t1 = "Stroma",
                    t2 = "Immune",
                    grouping_var = "group",
                    agg_fun = median)
```

If we instead want to look at how the distance between two cell types changes along a continuous variable, we can use the `plot_dist_scatter` function. The image-level data is again aggregated into patient-level data by a function `agg_fun`. Furthermore, we can plot a line of best fit using the argument `smooth`.

```{r}
mltplx_experiment %>%
  plot_dist_scatter(t1 = "Stroma",
                    t2 = "Immune",
                    cont_var = "age",
                    agg_fun = median,
                    smooth = "lm")
```

We can plot all of the distances between all pairs of cell types using the `plot_dist_matrix` function, which can be plotted as a network:

```{r}
mltplx_experiment %>%
  plot_dist_matrix(slide_ids = c("Slide 1", "Slide 2","Slide 3"),mode="network")
```

or a heatmap:

```{r}
mltplx_experiment %>%
  plot_dist_matrix(slide_ids = c("Slide 1", "Slide 2","Slide 3"),mode="heatmap")
```

Lastly, we can test for differences between distances in both patient groups for each cell type by using `lm_dist`. This function also allows us to adjust for patient-level variables, such as `age` in our example. The output from `lm_dist` can then be plotted with `plot_dist_regression_heatmap`.

```{r}
mltplx_experiment %>%
  lm_dist("group",
          covariates = c("age")) %>%
  plot_dist_regression_heatmap()
```


## Quantile-based analyses

The same sorts of analyses and plots can be generated for quantile-level analysis as well.

```{r}
mltplx_experiment %>%
  patient_boxplots_qdist(t1="Tumor",t2="Immune",grouping_var = "group")
```

```{r}
mltplx_experiment %>%
  plot_qdist_boxplots(t1 = "Stroma",
                    t2 = "Immune",
                    grouping_var = "group",
                    agg_fun = median)
```

```{r}
mltplx_experiment %>%
  plot_qdist_scatter(t1 = "Stroma",
                    t2 = "Immune",
                    cont_var = "age",
                    agg_fun = median,
                    smooth = "lm")
```

```{r}
mltplx_experiment %>%
  plot_qdist_matrix(slide_ids = c("Slide 1", "Slide 2","Slide 3"),mode="heatmap")
```

```{r}
mltplx_experiment %>%
  lm_qdist("group",
           interval = "67-100",
          covariates = c("age")) %>%
  plot_dist_regression_heatmap()
```

